\pdftitle

### 2018-05-14 maxNt_v3.tex
V2

%% Author: PGL  Porta Mana
%% Created: 2015-11-04T09:20:08+0100
%% Last-Updated: 2017-12-04T11:44:04+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Report-no: ***
\documentclass{article}

@@

\author{
  P.G.L. Porta Mana\\\texttt{mana@kth.org}
  \And
  V. Rostami\\\texttt{v.rostami@fz-juelich.de}
  \And
E. Torre\\\texttt{torre@ibk.baug.ethz.ch}
  % David S.~Hippocampus\thanks{Use footnote for providing further
  %   information about author (webpage, alternative
  %   address)---\emph{not} for acknowledging funding agencies.} \\
  % Department of Computer Science\\
  % Cranberry-Lemon University\\
  % Pittsburgh, PA 15213 \\
  % \texttt{hippo@cs.cranberry-lemon.edu} \\
  %% examples of more authors
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

@@

>>>>

%% Author: PGL  Porta Mana
%% Created: 2015-11-04T09:20:08+0100
%% Last-Updated: 2018-05-15T01:35:24+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Report-no: ***
\documentclass{article}

@@

\author{
  P.G.L. Porta Mana\\\texttt{piero.mana@ntnu.org}
  \And
  V. Rostami\\\texttt{v.rostami@fz-juelich.de}
  \AND
  E. Torre\\\texttt{torre@ibk.baug.ethz.ch}
    \And
Y. Roudi\\\texttt{yasser.roudi@ntnu.no}
  % David S.~Hippocampus\thanks{Use footnote for providing further
  %   information about author (webpage, alternative
  %   address)---\emph{not} for acknowledging funding agencies.} \\
  % Department of Computer Science\\
  % Cranberry-Lemon University\\
  % Pittsburgh, PA 15213 \\
  % \texttt{hippo@cs.cranberry-lemon.edu} \\
  %% examples of more authors
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

@@

Suppose that we have recorded the firing activity of a hundred neurons,
sampled from a particular brain area. This scenario is a concrete
possibility thanks to recent electrophysiological techniques
\cite{berenyietal2014}. In 1985, Gerstein, Perkel, \amp\ Dayhoff wrote (our
emphasis):
\begin{quote}
  the very profusion of data produced in such experiments creates serious
  problems of analysis and interpretation; these problems are conceptual as
  well as practical. The principal conceptual problems are (1)
  \emph{defining cooperativity or functional grouping} among neurons and
  (2) \emph{formulating quantitative criteria} for recognizing and
  characterizing such cooperativity. \cite{gersteinetal1985}
\end{quote}
The neuroscientific literature has since offered several mathematical
definitions of \enquote{cooperativity} or \enquote{functional grouping} and
criteria to quantify it.

One such quantitative criterion relies on the \me\ method
\cite{jaynes1957,jaynes1963,sivia1996_r2006,meadetal1984}. This criterion
has been used in neuroscience at least since the 1990s, applied to data
recorded from brain areas as diverse as retina and motor cortex
\cite{mackay1991,martignonetal1995,bohteetal2000,amarietal2003,schneidmanetal2006,shlensetal2006,mackeetal2009b,roudietal2009c,tkaciketal2009,gerwinnetal2009,mackeetal2011,mackeetal2011b,ganmoretal2011,granotatedgietal2013,tkaciketal2014b,moraetal2015,shimazakietal2015},
and it has been subjected to mathematical and conceptual scrutiny
\cite{tkaciketal2006,roudietal2009,roudietal2009b,barreiroetal2010b,mackeetal2013,rostamietal2016_r2017}.

\enquote{Cooperativity} can quantified and characterized with \me\ method
in several ways. The simplest way proceeds along the following steps.
Consider the recorded activity of a sample of $n$ neurons.
\begin{enumerate}
\item The activity of each neuron, ideally a continuous signal, is
  transformed into into a sequence of \enquote{$0$}s (inactive) and
  \enquote{$1$}s (firing) by binning the time and binarizing the signal
  intensity \cite{caianiello1961,caianiello1986}.

\item The total activity $n\yxs$ of the sample in each time bin is
  therefore an integer between $0$ and $n$. The total pairwise activity in
  each time bin, \ie\ the total number of simultaneously active pairs, is
  combinatorially given by $\tbinom{n\yxs}{2}\equiv n\yxs\,(n\yxs-1)/2$,
  and so on for higher-order simultaneous activities.

\item We construct a collection of \me\ distribution for the total
  activity, using as constraints: the time-averaged total activity; the
  time-averaged total activity \emph{and} total pairwise activity; and so
  on up to the time-averaged total activities of orders $1$ to $k$. Call
  the resulting distributions $p_1(n\yxs), p_2(n\yxs),\dotsc,p_k(n\yxs)$.
  We also have the empirical frequency distribution of the total activity,
  $f(n\yxs)$, counted from the time bins.

\item We can now compare the distributions above with one another and with
  the frequency distribution, using some probability-space distance like
  the relative entropy or discrimination information
  \citep{kullback1987,jaynes1963,hobson1969,hobsonetal1973}. If we find,
  for example, that such distance is very high between $p_1$ and $f$, but
  very low between $p_k$ and $f$ for $k \ge 2$, then we can say that there
  is a \enquote{pairwise cooperativity}, and that any higher-order
  cooperativity is just a reflection or consequence of the pairwise one.
\end{enumerate}
The protocol above needs to be made precise by specifying various
parameters, for example the width of the time bins.

We hurry to say that the description just given is not \emph{the} way to
quantify and characterize cooperativity and functional grouping, but just
\emph{one} way. It can surely be criticized from many points of view. Yet,
it is quantitative and more precise than an undefined notion of
\enquote{cooperativity}. Two people who apply it to the same data will
obtain the same numbers. Other protocols can be based on the \me\ method,
for example protocols that take into account the activities or joint
activities of individual neurons.

The purpose of the present work is to show that there is a problem in this
general way of applying the \me\ method to the data recorded from our
sampled neurons.

Suppose that the neurons were recorded with electrodes covering an area of
some square millimetres. This recording is a \emph{sample} of the activity
of the population covered by the electrode; or we can consider it as a
sample of an even more extended brain area.

The characterization of the cooperativity of the recorded sample would have
little meaning if we did not expect its results to generalize to a larger
area, at least that covered by the electrodes. In other words, we expect
that the conclusions drawn with \me\ methods about the sampled neurons
should somehow generalize or extrapolate at least to larger areas from
which those neurons were sampled. We are thus assuming that the recorded
sample is a \emph{representative sample} of some larger area.

The problem is this. We can apply the \me\ method to the sample, as
described in the above protocol, to generate probability distributions for
the state of the sample. But, if our sample is representative of a larger
population, we can apply the \me\ method to the larger population, using
the constraints obtained from the sample, and generate a probability
distribution for the larger population; the distribution for the sample is
then obtained by marginalization. Either application seems to have some
commendable features. However, \emph{the distributions obtained from these
  two applications differ}. Which choice is most meaningful?
